### 2014-03-17 msc-thesis/chapters/1/chapter-intro.tex
MSC final melanie edits.

%Describe difference between direct and indirect
Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts are caused by source code changes that negatively impact another
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influence negatively. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. However,
previous interviews and surveys conducted with software developers have shown a pattern that developers 
of a software project view indirect conflict awareness  as a high priority issue in their 
development~\cite{Damian:2007:GSE, Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}, meaning that future research
is required to address this developer concern.

@@

% What is being addressed in this thesis
Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
with supporting indirect conflicts through awareness techniques as well as to explore possible solution for industry practice 
in the are of indirect conflicts. To accomplish this goal, I have researched the following sub topics of indirect
conflicts: technical dependencies, how developers are involved in said dependencies,
socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
are being used, what future steps should be taken by researchers to better industry regarding indirect conflicts, and finally,
how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
conducting the following four studies.

@@

% Study 1 introduction
Study 1 focuses on the power of technical dependencies in software projects. The question I
investigated were : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
code changes statistically relate to bugs?}''. This study explains the approach used to locate these pairs of developers 
in developer networks.The process utilizes code changes and the call hierarchies effected to find patterns of developer 
relationships in successful and failed code changes. These developers pairs can be seen as indirect conflicts occurring
as one developer's code change has negatively affects another developer's work. As it will be seen, I found 27 statistically significant failure 
inducing developer pairs. These developer relationships can be used to promote the idea of leveraging socio-technical 
congruence, a measure of coordination compared to technical dependencies amongst stakeholders, to provide coordination recommendations.
This notion of socio-technical congruence is my initial proposed solution to indirect conflicts. By identifying these failure
inducing pair of developers over indirect conflicts, I hoped that recommended communication would be the correct fix. The 
results of Study 1 directly influence Study 2.

% Study 2 introduction
Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool while
answering: ``\textit{Can indirect conflicts be supported through an awareness mechanism which leverages pairs of
developers who's changing technical dependencies statistically relate to bugs?}''.
In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
evaluation of the tool \textit{Impact}, a web based tool that aims at detecting indirect conflicts among developers and 
notifying the appropriate members involved in these conflicts. By leveraging technical relationships inherent of software 
projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationship 
through software configuration management (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as 
alert developers involved in such conflicts in task awareness. While this study outlines \textit{Impact's} specific 
implementation, its design is rather generic and can be implemented in similar indirect conflict awareness tools. 
\textit{Impact} represents a first step towards the design and implementation of awareness tools which address indirect 
conflicts in software development. After a brief evaluation of \textit{Impact} with two student software teams, it was
found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
large problems found in many other indirect conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to fully understand 
the causes of these indirect conflict tool issues, a third study was conducted.

@@

I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. The study findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of those changes,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts. As a result of the final finding (the gap in software evolution analytical tools), I conducted
a fourth and final study.

@@

% The rest of the thesis 
The remainder of this dissertation is laid out as follows. Chapter two includes Study 1 and Study 2 as the motivational studies which
ultimately led to the larger research en devours found in Study 3 and Study 4. Chapter 3 includes Study 3 and Study 4 which ultimately
press upon indirect conflicts in a more in depth fashion than has been previously seen in research. Chapter 4 contains a lengthy discussion
of what has been learned from all four studies of this dissertation as well as implications for further research and tool development
in the field of indirect conflicts. Finally, Chapter 5 concludes this dissertation.

>>>>

%Describe difference between direct and indirect
Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts are caused by source code changes that negatively impact another
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influenced negatively. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. However,
previous interviews and surveys conducted with software developers have shown a pattern that developers 
of a software project view indirect conflict awareness  as a high priority issue in their 
development~\cite{Damian:2007:GSE, Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}, meaning that future research
is required to address this developer concern.

@@

% What is being addressed in this thesis
Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
with supporting indirect conflicts through awareness techniques as well as to explore possible solutions for industry practice 
in the area of indirect conflicts. To accomplish this goal, I have researched the following sub topics of indirect
conflicts: technical dependencies, how developers are involved in said dependencies,
socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
are being used, what future steps should be taken by researchers to better industry regarding indirect conflicts, and finally,
how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
conducting four studies.

@@

% Study 1 introduction
Study 1 focuses on the power of technical dependencies in software projects. The question I
investigated were : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
code changes statistically relate to bugs?}''. This study explains the approach used to locate these pairs of developers 
in developer networks.The process utilizes code changes and the call hierarchies effected to find patterns of developer 
relationships in successful and failed code changes. These developer pairs can be seen as indirect conflicts occurring
as one developer's code change has negatively affects another developer's work. As it will be seen, I found 27 statistically significant failure 
inducing developer pairs. These developer relationships can be used to promote the idea of leveraging socio-technical 
congruence, a measure of coordination compared to technical dependencies amongst stakeholders, to provide coordination recommendations.
This notion of socio-technical congruence is my initial proposed solution to indirect conflicts. By identifying these failure
inducing pair of developers over indirect conflicts, I hoped that recommended communication would be the correct fix. The 
results of Study 1 directly influence Study 2.

% Study 2 introduction
Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool while
answering: ``\textit{Can indirect conflicts be supported through an awareness mechanism which leverages pairs of
developers whose changing technical dependencies statistically relate to bugs?}''.
In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
evaluation of the tool \textit{Impact}, a web based tool that aims at detecting indirect conflicts among developers and 
notifying the appropriate members involved in these conflicts. By leveraging technical relationships inherent of software 
projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationships 
through software configuration management (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as 
alert developers involved in such conflicts in task awareness. While this study outlines \textit{Impact's} specific 
implementation, its design is rather generic and can be implemented in similar indirect conflict awareness tools. 
\textit{Impact} represents a first step towards the design and implementation of awareness tools which address indirect 
conflicts in software development. After a brief evaluation of \textit{Impact} with two student software teams, it was
found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
large problems found in many other indirect conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to fully understand 
the causes of these indirect conflict tool issues, a third study was conducted.

@@

I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. The study findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of those changes,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and that there exists a gap in software evolution analytical tools arising from the reliance on static analysis resulting in missed
context of indirect conflicts. As a result of the final finding (the gap in software evolution analytical tools), I conducted
a fourth and final study.

@@

% The rest of the thesis 
The remainder of this dissertation is laid out as follows. Chapter two includes Study 1 and Study 2 as the motivational studies which
ultimately led to the larger research studies found in Study 3 and Study 4. Chapter 3 includes Study 3 and Study 4 which ultimately
press upon indirect conflicts in a more in depth fashion than has been previously seen in research. Chapter 4 contains a lengthy discussion
of what has been learned from all four studies of this dissertation as well as implications for further research and tool development
in the field of indirect conflicts. Finally, Chapter 5 concludes this dissertation.

### 2014-03-17 msc-thesis/chapters/2/chapter-motivating.tex
MSC final melanie edits.

The second study, ``Awareness with Impact'' ((Section~\ref{study:impact})), 
takes the notion of developer
pairs in indirect conflicts learned from Study 1, and adds in source code change detection in order to create
an awareness notification system for developers called \textit{Impact}. \textit{Impact} was designed to a developer
to any source code changes preformed by another developers when the two are linked in a technical dependency through
a developer pair. \textit{Impact} utilized a non-obtrusive RSS style feed for notifications for visualization. 
While \textit{Impact} showed some promise through its user evaluation, it ultimately suffered the fate of information
overload as was seen in other indirect conflict tools~\cite{Sarma:2007:TSA,Servant:2010:CPI,Trainer:2005:BGT}.

>>>>

The second study, ``Awareness with Impact'' ((Section~\ref{study:impact})), 
takes the notion of developer
pairs in indirect conflicts learned from Study 1, and adds in source code change detection in order to create
an awareness notification system for developers called \textit{Impact}. \textit{Impact} was designed to alert a developer
to any source code changes preformed by another developer when the two are linked in a technical dependency through
a developer pair. \textit{Impact} utilized a non-obtrusive RSS style feed for notifications. 
While \textit{Impact} showed some promise through its user evaluation, it ultimately suffered the fate of information
overload as was seen in other indirect conflict tools~\cite{Sarma:2007:TSA,Servant:2010:CPI,Trainer:2005:BGT}.

### 2014-03-17 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC final melanie edits.

% Explain technical network research that already exists
Technical dependencies in a project can be used to predict success or failure in
builds or code changes~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU}. However, most 
research in this area is based on identifying
central modules inside a large code base which are likely to cause software failures or
detecting frequently changed code that can be associated with previous failures
~\cite{Kim:2006:AIB}. 
This module-based method also results in predictions 
at the file or binary level of software development as opposed to a code change level
and often lack the ability to provide recommendations for improved coordination
other than test focus.

@@

On the other side with the human factor, researchers have examined predicting build outcomes of software
using communication patterns from developers. Wolf et al.~\cite{5070503} used patterns of communication from between
pre-existing build in order to predict later build outcomes. Naggappan et al.~\cite{nagappan08} showed that having a large
communication organization difference between developers who worked on the same software module had
a negative influence on the quality of the software.

@@

% Changeset effects
The code change effect, if any, to the call hierarchy is now found. The Git
software repository is used to determine what changes were made to any give file inside a 
code change. Specifically, methods modified by a code change are searched for. The call graph 
is then used to determine which methods call those that have been changed, which
gives the code change technical dependencies.

>>>>

% Explain technical network research that already exists
Technical dependencies in a project can be used to predict success or failure in
builds or code changes~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU}. However, most 
research in this area is based on identifying
central modules inside a large code base which are likely to cause software failures or
detecting frequently changed code that can be associated with previous failures
~\cite{Kim:2006:AIB}. 
These module-based methods also result in predictions 
at the file or binary level of software development as opposed to a code change level
and often lack the ability to provide recommendations for improved coordination
other than test focus.

@@

On the other side with the human factor, researchers have examined predicting build outcomes of software
using communication patterns from developers. Wolf et al.~\cite{5070503} used patterns of communication from between
pre-existing builds in order to predict later build outcomes. Naggappan et al.~\cite{nagappan08} showed that having a large
communication organizational difference between developers who worked on the same software module had
a negative influence on the quality of the software.

@@

% Changeset effects
The code change effect, if any, to the call hierarchy is now found. The Git
software repository is used to determine what changes were made to any given file inside a 
code change. Specifically, methods modified by a code change are searched for. The call graph 
is then used to determine which methods call those that have been changed, which
gives the code change technical dependencies.

### 2014-03-17 msc-thesis/chapters/2/sec-impact.tex
MSC final melanie edits.

%Outline of study
In this study, I report on research into supporting developer pairs in indirect conflicts
and present the design, implementation, and future evaluation of the tool \textit{Impact},
a web based tool that aims at detecting indirect conflicts among developers
and notifying the appropriate members involved in these conflicts. Through Impact and
its evaluation I ask:

\begin{description}
        \item[RQ\namedlabel{itm:rq1-impact}{RQ1-Impact}] \textit{Can indirect conflicts be supported 
        through an awareness mechanism which leverages pairs of developers who's changing technical 
        dependencies statistically relate to bugs?}
\end{description}

@@

I first start with both Codebook~\cite{Begel:2010:CDE} and 
Ariadne~\cite{Trainer:2005:BGT}. These projects produce an exploratory
environment for developers to handle indirect conflicts. Exploratory
pertains to the ability to solve self determined conflicts, meaning that
once a developer discovers a conflict, they can use the tool as a type of
lookup table to solve their issue. Codebook is a type of social developer
network that relates developers to source code, issue repositories and
other social media while Ariadne only examines there source code for developer
to source code association. Through Codebook, developers become
owners of source code artifacts. Both projects also use program 
dependency graphs~\cite{Horwitz:1992:UPD}
in order to relate technical artifacts to each other. These projects make 
use of method call graphs in order to 
determine which methods invoke others which forms the basis for 
linking source code artifacts creating a directed graph. While these 
projects can be great tools 
for solving indirect conflicts which may arise, by querying such directed
graphs to view impacts of conflict creating code, they lack the ability to
detect potential conflicts on their own.

A serious attempt at both detecting and informing developers of
indirect conflicts is the tool Palantir~\cite{Sarma:2007:TSA}. Palantir
monitors developers activities in files with regards to class signatures.
Once a developer changes the signature of a class, such as by modifying changes
in name, parameters, or return values of public methods, any workspace
of other developers which are using that class will be notified. Palantir utilizes
a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
a favored collection system among awareness tools. Sarma et al.
~\cite{Sarma:2007:TSA} also
developed a generic design for future indirect conflict awareness tools. 
However, Palantir falls short in its collection and distribution
mechanisms. First, Palantir only considers ``outside'' appearance of technical
objects, being their return types, parameters, etc. Secondly, Palantir 
only delivers
detected conflicts to developers who are presently viewing or editing
the indirect object while other developers who have used the modified 
class previously are not notified.

I will lastly examine the tool CASI~\cite{Servant:2010:CPI} which uses
a sphere of influence for each developer to determine how source code
changes are indirectly related to other components of a software project.
CASI uses dependency slicing~\cite{Bajracharya:2009:SIS} instead of the 
call graphs in Ariadne~\cite{Trainer:2005:BGT} which gives dependencies among
all source code entities. This provides a verbose output of dependencies when
source code is changed. CASI also implements a visualization where
a developer can see what parts of a software projects he or she may be effecting
with the source code change. This allows the developer themselves to go and fix
potential issues elsewhere in the project before the code change is committed
to the software repository. While CASI covers great ground in its approach,
it still leaves the issue of information overload, although attempts were made
to solve this by having severity levels of indirect conflicts presented to
the user.

@@

Finally, the indirect conflict tool must be able to detect changes
to the technical objects defined above and notify the appropriate owners
to the conflict. 
Two approaches have been 
proposed for change gathering techniques: real time and commit 
time~\cite{Fitzpatrick:2002:SPA}.
I propose the use of commit time
information gathering as it avoids the issue of developers 
overwriting previous work or deleting modifications which would 
produce information for changes that no longer exist. However, the
trade off is that indirect conflicts must be committed before detected,
which results in conflicts being apart of the system before being able
to be dealt with as opposed to catching conflicts before they happen.
At commit time, the tool must parse changed source code in relation
to technical artifacts in the created directed graph detailed above.
Where \textit{Impact's} design differs from that of Palantir's is that
the object's entire body (method definition and internal body) is 
parsed, similar to that of CASI~\cite{Servant:2010:CPI},
at commit time, as opposed to real time, to detect changes anywhere in the technical object.
This is a first design step towards avoiding information overload.
Once technical objects are found to be changed, appropriate owners
of objects which use the changed object should be notified. 
In Figure~\ref{fig:graph}, Carl changes method (technical object) 1,
which effects methods 2 and 3 resulting in the alerting of
developers Bret, Daniel, and Anne. I have opted to alert the invoking
developers rather than the developer making the change to potential
solutions as my conflicts are detected at commit time and this supports
the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} 
from software structure to communication patterns in awareness systems.

@@

I asked lead developers to address two main concerns: do indirect
conflicts pose a threat
at the method level (e.g. method 1 has a bug because it invokes method
2 which has had its implementation changed recently), 
and did \textit{Impact} help raise
awareness and promote quicker conflict resolution for indirect
conflicts. The two interviews largely supported the expectation of
indirect conflicts posing a serious threat to developers, especially
in medium to large teams or projects as opposed to the small
teams which they were apart of. It was also pointed
out that method use can be a particularly large area for indirect
conflicts to arise. However, it was noted that
any technical object which is used as an interface to some data
construct or methodology, database access for instance, can be 
a large potential issue for indirect conflicts.  Interview responses to
\textit{Impact} were optimistically positive, as interviewees stated that \textit{Impact}
had potential to raise awareness among their teams with what other developers
are doing as well as the influence it has on their own work. However,
\textit{Impact} was shown to have a major problem of information overload. 
It was suggested
that while all method changes were being detected,
not all are notification worthy. One developer suggested to only notify
developers to indirect conflicts if the internal structure of a method
changes due to modification to input parameters or output parameters.
In other words, the boundaries of the technical objects (changing
how a parameter is used inside the method, modifying the return
result inside the method) seem to be more of interest than other 
internal workings. More complex inner workings of methods were also noted 
to be of interest to developers
such as cyclomatic complexity, or time and space requirements.

@@

\subsection{Threats to Validity}
Because of the tool validation nature of this work, I chose participant interviews as my research validation method, which has some implications regarding the
limitations and threats to validity of this study. While I did some positive results regarding the potential of Impact in this study
populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.

As per my interview participants, I conducted this case study with undergraduate students at the University of Victoria. This being said, participants
may not have had enough real world experience to validate this study at an industry level. The students were also consumed with their regular course
work which could have limited the time spend using Impact and the enthusiasm put forward while conducting this study.

To counter act this, my study was conducted purely on a volunteer basis to eliminate those participants which may have been to busy to focus
on the study or provide appropriate feedback where needed.

>>>>

%Outline of study
In this study, I report on research into supporting developer pairs in indirect conflicts
and present the design, and implementation of the tool \textit{Impact},
a web based tool that aims at detecting indirect conflicts among developers
and notifying the appropriate members involved in these conflicts. Through Impact and
its evaluation I ask:

\begin{description}
        \item[RQ\namedlabel{itm:rq1-impact}{RQ1-Impact}] \textit{Can indirect conflicts be supported 
        through an awareness mechanism which leverages pairs of developers whose changing technical 
        dependencies statistically relate to bugs?}
\end{description}

@@

I first start with both Codebook~\cite{Begel:2010:CDE} and 
Ariadne~\cite{Trainer:2005:BGT}. These projects produce an exploratory
environment for developers to handle indirect conflicts. Exploratory
pertains to the ability to solve self determined conflicts, meaning that
once a developer discovers a conflict, they can use the tool as a type of
lookup table to solve their issue. Codebook is a type of social developer
network that relates developers to source code, issue repositories and
other social media while Ariadne only examines their source code for developer
to source code association. Through Codebook, developers become
owners of source code artifacts. Both projects also use program 
dependency graphs~\cite{Horwitz:1992:UPD}
in order to relate technical artifacts to each other. These projects make 
use of method call graphs in order to 
determine which methods invoke others which forms the basis for 
linking source code artifacts creating a directed graph. While these 
projects can be great tools 
for solving indirect conflicts which may arise, by querying such directed
graphs to view impacts of conflict creating code, they lack the ability to
detect potential conflicts on their own.

A serious attempt at both detecting and informing developers of
indirect conflicts is the tool Palantir~\cite{Sarma:2007:TSA}. Palantir
monitors developer's activities in files with regards to class signatures.
Once a developer changes the signature of a class, such as by modifying changes
in name, parameters, or return values of public methods, any workspace
of other developers which are using that class will be notified. Palantir utilizes
a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
a favored collection system among awareness tools. Sarma et al.
~\cite{Sarma:2007:TSA} also
developed a generic design for future indirect conflict awareness tools. 
However, Palantir falls short in its collection and distribution
mechanisms. First, Palantir only considers ``outside'' appearance of technical
objects, being their return types, parameters, etc. Secondly, Palantir 
only delivers
detected conflicts to developers who are presently viewing or editing
the indirect object while other developers who have used the modified 
class previously are not notified.

I will lastly examine the tool CASI~\cite{Servant:2010:CPI} which uses
a sphere of influence for each developer to determine how source code
changes are indirectly related to other components of the software.
CASI uses dependency slicing~\cite{Bajracharya:2009:SIS} instead of the 
call graphs in Ariadne~\cite{Trainer:2005:BGT} which gives dependencies among
all source code entities. This provides a verbose output of dependencies when
source code is changed. CASI also implements a visualization where
a developer can see what parts of a software projects he or she may be effecting
with the source code change. This allows the developers themselves to go and fix
potential issues elsewhere in the project before the code change is committed
to the software repository. While CASI covers great ground in its approach,
it still leaves the issue of information overload, although attempts were made
to solve this by having severity levels of indirect conflicts presented to
the user.

@@

Finally, the indirect conflict tool must be able to detect changes
to the technical objects defined above and notify the appropriate owners
to the conflict. 
Two approaches have been 
proposed for change gathering techniques: real time and commit 
time~\cite{Fitzpatrick:2002:SPA}.
I propose the use of commit time
information gathering as it avoids the issue of developers 
overwriting previous work or deleting modifications which would 
produce information for changes that no longer exist. However, the
trade off is that indirect conflicts must be committed before detected,
which results in conflicts being integrated into the system before being able
to be dealt with as opposed to catching conflicts before they happen.
At commit time, the tool must parse changed source code in relation
to technical artifacts in the created directed graph detailed above.
Where \textit{Impact's} design differs from that of Palantir's is that
the object's entire body (method definition and internal body) is 
parsed, similar to that of CASI~\cite{Servant:2010:CPI},
at commit time, as opposed to real time, to detect changes anywhere in the technical object.
This is a first design step towards avoiding information overload.
Once technical objects are found to be changed, appropriate owners
of objects which use the changed object should be notified. 
In Figure~\ref{fig:graph}, Carl changes method (technical object) 1,
which effects methods 2 and 3 resulting in the alerting of
developers Bret, Daniel, and Anne. I have opted to alert the invoking
developers rather than the developer making the change to potential
solutions as my conflicts are detected at commit time and this supports
the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} 
from software structure to communication patterns in awareness systems.

@@

I asked lead developers to address two main concerns: do indirect
conflicts pose a threat
at the method level (e.g. method 1 has a bug because it invokes method
2 which has had its implementation changed recently), 
and did \textit{Impact} help raise
awareness and promote quicker conflict resolution for indirect
conflicts. The two interviews largely supported the expectation of
indirect conflicts posing a serious threat to developers, especially
in medium to large teams or projects as opposed to the small
teams which they were a part of. It was also pointed
out that method use can be a particularly large area for indirect
conflicts to arise. However, it was noted that
any technical object which is used as an interface to some data
construct or methodology, database access for instance, can be 
a large potential issue for indirect conflicts.  Interview responses to
\textit{Impact} were optimistically positive, as interviewees stated that \textit{Impact}
had potential to raise awareness among their teams with what other developers
are doing as well as the influence it has on their own work. However,
\textit{Impact} was shown to have a major problem of information overload. 
It was suggested
that while all method changes were being detected,
not all are notification worthy. One developer suggested to only notify
developers if the internal structure of a method
changes due to modification to input parameters or output parameters.
In other words, the boundaries of the technical objects (changing
how a parameter is used inside the method, modifying the return
result inside the method) seem to be more of interest than other 
internal workings. More complex inner workings of methods were also noted 
to be of interest to developers
such as cyclomatic complexity, or time and space requirements.

@@

\subsection{Threats to Validity}
Because of the tool validation nature of this work, I chose participant interviews as my research validation method, which has some implications regarding the
limitations and threats to validity of this study. While I did have some positive results regarding the potential of Impact in this study,
populations studied from outside of this study's participants may add new insights into the pool of findings. As a result of this, findings
from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.

I conducted this case study with undergraduate students at the University of Victoria. This being said, participants
may not have had enough real world experience to validate this study at an industry level. The students were also consumed with their regular course
work which could have limited the time spent using Impact and the enthusiasm put forward while conducting this study.

To counter this, my study was conducted purely on a volunteer basis to eliminate those participants which may have been to busy to focus
on the study or provide appropriate feedback where needed.

### 2014-03-17 msc-thesis/chapters/3/chapter-exploration.tex
MSC final melanie edits.

Through the two previous studies, I have shown that developers linked indirectly in source code changes can statistically
related to software failures. In the attempts of mitigating these loses through added awareness, I implemented
an indirect conflict tool called \textit{Impact}. However, \textit{Impact} ultimately suffered from information
overload as seen in its evaluation which was caused by false positives and scalability of the tool.

@@

Based on some of the findings (to be seen in details in Section~\ref{study:exploration}) I performed a follow up study
which did not relate directly to the themes of this dissertation but helped strengthen the results found in Study 3.
Some results of Study 3 showed that indirect conflict tools should take into account a contextual setting of development
progression in software projects to better inform developers of potential indirect conflicts. In other words, an indirect
conflict tool should be able to tell what phase of development inside the development life cycle a project is currently
active in. In order to better explore and support this finding, I performed a complimentary study of software evolutionary trends (Section~\ref{study:apie}).
In Study 4, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near major release and more
frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.

>>>>

Through the two previous studies, I have shown that developers linked indirectly in source code changes can be statistically
related to software failures. In the attempts of mitigating these loses through added awareness, I implemented
an indirect conflict tool called \textit{Impact}. However, \textit{Impact} ultimately suffered from information
overload as seen in its evaluation which was caused by false positives and scalability of the tool.

@@

Based on some of the findings (to be seen in details in Section~\ref{study:exploration}) I performed a follow up study
which did not relate directly to the themes of this dissertation but helped strengthen the results found in Study 3.
Some results of Study 3 showed that indirect conflict tools should take into account a contextual setting of development
progression in software projects to better inform developers of potential indirect conflicts. In other words, an indirect
conflict tool should be able to tell what phase of development inside the development life cycle a project is currently
active in. In order to better explore and support this finding, I performed a complimentary study of software evolutionary trends (Section~\ref{study:apie}).
In Study 4, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near major release and more
frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
to determine the probability of indirect conflicts occurring and thus aiding developers in dealing with indirect conflicts in their projects.

### 2014-03-17 msc-thesis/chapters/3/sec-apie.tex
MSC final melanie edits.

While measuring software quality has had a major focus in software engineering research for many years~\cite{Bowen:1978:CAS}~\cite{Grady:1993:PRM}~\cite{ISOIEC9126},
the sub study of software stability and its implications on reliability and indirect conflict proneness remains a difficult subject to understand. 
The decision of what makes a project stable and ready
for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
the project~\cite{Conway:1968}. Code churn is an often looked to statistically for stability but can be grossly misleading
in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC} as well as proneness to indirect conflicts
both internally and externally to third party applications. 
Creating an approach to determining software stability, release preparedness, and the proneness of indirect conflicts is still a large open area of
interest in software engineering research.

@@

Once the ChangeDistiller information was collected, I decided to examine software change trends surrounding releases of the project's I had selected. Since releases
have preconceived notions of software stability and a lack of proneness to indirect conflicts, 
I decided that by studying the change types surrounding these releases, I could get a better understanding of
what types of source code changes or trends constitute software stability or maturity. In order to study the release points, I went to each of my 10 project's 
web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total I identified 472 releases
across my 10 studied projects.

@@

As it can be seen in Table~\ref{tab:apie-ratio}, there are few change type trends around major releases which pass my threshold. We can see that both public
and private methods
being removed from a project is more likely to occur after a major release than after. Table~\ref{tab:apie-ratio} also shows significance in the changes to private
classes. We see that private classes are added more (24\%) before a major release and removed more after (44\%) the release. 
All results in Table~\ref{tab:apie-ratio} could be used as identified trends of major software releases, while I have just highlighted the larger ratios
which meet my threshold criteria.

@@

These 9 change trends can be used in future indirect conflict tools in order to identify a context for indirect conflicts.
For example. Any of the 9 change trends which occur more so after a major release may be used as a sign of high indirect conflict proneness
since after a major release a new development cycle is likely to begin. As per change trends which occur more so before a major release, indirect
conflicts may use this context in order to identify a low proneness to indirect conflicts in their results. These two contextual patterns
can be used throughout the life of a software project in order to help better inform indirect conflict tools as to the processes of developers
and provide better feedback to said developers of indirect conflicts.

>>>>

While measuring software quality has had a major focus in software engineering research for many years~\cite{Bowen:1978:CAS}~\cite{Grady:1993:PRM}~\cite{ISOIEC9126},
the study of software stability and its implications on reliability and indirect conflict proneness remains a difficult subject to understand. 
The decision of what makes a project stable and ready
for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
the project~\cite{Conway:1968}. Code churn is an often looked to statistically for stability but can be grossly misleading
in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC} as well as proneness to indirect conflicts
both internally and externally to third party applications. 
Creating an approach to determining software stability, release preparedness, and the proneness of indirect conflicts is still a large open area of
interest in software engineering research.

@@

Once the ChangeDistiller information was collected, I decided to examine software change trends surrounding releases of the projects I had selected. Since releases
have preconceived notions of software stability and a lack of proneness to indirect conflicts, 
I decided that by studying the change types surrounding these releases, I could get a better understanding of
what types of source code changes or trends constitute software stability or maturity. In order to study the release points, I went to each of my 10 project's 
web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total I identified 472 releases
across my 10 studied projects.

@@

As it can be seen in Table~\ref{tab:apie-ratio}, there are few change type trends around major releases which pass my threshold. We can see that both public
and private methods
being removed from a project is more likely to occur after a major release than before. Table~\ref{tab:apie-ratio} also shows significance in the changes to private
classes. We see that private classes are added more (24\%) before a major release and removed more after (44\%) the release. 
All results in Table~\ref{tab:apie-ratio} could be used as identified trends of major software releases, while I have just highlighted the larger ratios
which meet my threshold criteria.

@@

These 9 change trends can be used in future indirect conflict tools in order to identify a context for indirect conflicts.
For example. Any of the 9 change trends which occur more so after a major release may be used as a sign of high indirect conflict proneness
since after a major release a new development cycle is likely to begin. As per change trends which occur more so before a major release, indirect
conflict tools may use this context in order to identify a low proneness to indirect conflicts in their results. These two contextual patterns
can be used throughout the life of a software project in order to help better inform indirect conflict tools as to the processes of developers
and provide better feedback to said developers of indirect conflicts.
